---
title: "Tx Import for DESeq2"
author: "Chelsea Herdman"
date: "30/01/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In order to perform differential expression analysis using DESeq2 on the 
estimated transcript abundances produced by kallisto, we used the tximport
package in R.

We followed the vignette provided by Michael I. Love, Charlotte Soneson and 
Mark D. Robinson found [here](http://bioconductor.org/packages/release/bioc/vignettes/tximport/inst/doc/tximport.html#salmon_with_inferential_replicates)


## Prepare transcript to gene dataframe for gene-level summarization

Load required libraries.

```{r libraries, message=FALSE, warning=FALSE}
library(tximport)
library(data.table)
library(ensembldb)
library(RMariaDB)
library(DESeq2)
```

Fetch gene annotations from Ensembl and create one to one transcript id/gene id data frame.

```{r tx2gene}
txdb = makeTxDbFromEnsembl(organism="Danio rerio",
                    release=99,
                    circ_seqs=DEFAULT_CIRC_SEQS,
                    server="ensembldb.ensembl.org",
                    username="anonymous", password=NULL, port=0L,
                    tx_attrib=NULL)

k <- keys(txdb, keytype = "TXNAME")
tx2gene <- select(txdb, k, "GENEID", "TXNAME")
head(tx2gene)

length(unique(tx2gene$TXNAME))

length(unique(tx2gene$GENEID))

```

## Prepare named vector of files

Use kallisto abundance.h5 files. These are listed in alphabetical order in the 
kallisto_out directory so ensure naming associates accurately and that names 
associate with sample info table that will be used for DESeq2.

```{r makefilesvec}

files = list.files(path="../../kallisto_out/",
                   pattern="abundance.h5",
                   recursive = TRUE,
                   full.names = TRUE)

all(file.exists(files))
#[1] TRUE

names(files) = paste0("14893X", c(1, 10:19, 2, 20:22, 3:9))

files
```

## Run tximport for gene level estimation
```{r tximport}
txi.kallisto = tximport(files, type= "kallisto", tx2gene = tx2gene, ignoreTxVersion = TRUE)
head(txi.kallisto$counts)
```

***

## Create DESeq DataSet

As described in the vignette, this function creates an offset in order to perform differential expression analysis.
Will also run with countsFromAbundance="lengthScaledTPM" and compare
results for counts histograms etc.
```{r dds}
sampleTable = fread("../../ribozero_sample_info.txt")

cData = data.frame(time_point=factor(sampleTable$time_point,
                                     levels=c("24hpf", "36hpf", "48hpf", "60hpf",
                                              "72hpf")),
                   replicate_id=factor(sampleTable$rep_id,
                                       levels=paste("rep_", 1:5, sep="")))
                                              
rownames(cData) = sampleTable$gnomex_id

dds <- DESeqDataSetFromTximport(txi.kallisto, cData, ~time_point)
```

***
## References
* Bray N, Pimentel H, Melsted P, Pachter L (2016), _Near-optimal probabilistic RNA-seq quantification_, Nature Biotechnology, 34, 525–527. doi:10.1038/nbt.3519  
* Soneson C, Love MI, Robinson MD (2015), _Differential analyses for RNA-seq: transcript-level estimates improve gene-level inferences_, F1000Research, 4, 1521. doi:10.12688/f1000research.7563.2  
* Love MI, Huber W, Anders S (2014), _Moderated estimation of fold change and dispersion for RNA-seq data with DESeq2_, Genome Biology, 15, 550. doi:10.1186/s13059-014-0550-8  

***
